---
title: "exercise 5.10"
author: "Sabrina Garcia"
date: "May 17, 2018"
output: html_document
---

# First Header

Exercise 5.10 in reproducible r workshop

## Second Header

Text!

### Third Header

#### Fourth Header

Make sure you pull, commit, pull, push!

# Tidy Data

``` {r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE)
```

``` {r Libraries}
library(dplyr)
library(tidyr)
```

Reads in data from KNB [here](https://knb.ecoinformatics.org/#view/df35b.304.2)

``` {r read_in_data}
catch_df <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1", method = "libcurl"),
                    stringsAsFactors = FALSE)
head(catch_df)
```

```{r unique}
unique(catch_df$notesRegCode)
```

Here we selected the columns we actually want

``` {r remove_columns}
catch_df <- catch_df %>% 
  select(Region, Year, Chinook, Sockeye, Coho, Pink, Chum)
```

You can also exclude column names by including a minus sign in front
of the column name select(-All, -notesRegCode)

Still considered wide data, we can add a species column to make it tidy

In gather, the key is the new column name and the values are the cell values
You have to tell gather what to ignore

``` {r gather}
catch_df <- catch_df %>% 
  gather( key = species, value=catch, -Region, -Year )
```

Spread does the opposite of gather and makes your data into wide format

```{r spread}
test_spread <- catch_df %>% 
  spread(key=species, value=catch)
```

```{r rename_columns}
catch_df <- catch_df %>% 
  rename( catch_thousands = catch, Species = species)
```

Correct typo and change the units of catch column.

```{r change_value_in_data}
catch_df <- catch_df %>% 
  mutate( catch_thousands = ifelse( catch_thousands == "I", 1, catch_thousands )) %>% 
  mutate( catch_thousands = as.integer(catch_thousands) ) %>% 
  mutate( catch = catch_thousands * 1000 )  
```

This is a good place to commit because we just finished a functional block of code.




